Parallelization of Fast $\ell_1$-Minimization for Face Recognition

### 2011-03-01 paper/main.tex
Fixed compile bug; hyperref needs to be uspackaged before algorithm

\usepackage{iccv}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Packages added by us
\usepackage{amsfonts}
\usepackage{mydefs}
\usepackage{algorithmic}
\usepackage{algorithm}

\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

@@

One of the drawbacks of most interior-point methods for $\ell_1$-min (including
BP, LASSO, and TNIPM) is that they require the solution sequence to follow an
interior path{\bf DW: Following the interior path is what made them faster than
traditional linear programming techniques that follow the edge of the feasible set;
does Homotopy actually go outside the feasible set?  Just want to check if this is
an accurate representation...}, which is often computationally expensive. To mitigate these
issues, an approach called \emph{Homotopy} has been studied to accelerate the
speed of $\ell_1$-min. Homotopy methods for recovering sparse signals were
first studied in the context of LASSO \cite{OsborneM2000}, which inspired a
solution to the \emph{forward stagewise linear regression} problem called LARS
\cite{EfronB2004}, and eventually led to the corresponding Homotopy algorithms
for BP in \cite{MalioutovD2005,DonohoD2006}.

@@

Note that in the iterative procedure outlined in \eqref{eq:alm}, the second
step only involves vector algebra and matrix-vector multiplication. Therefore,
the procedure is computationally efficient only if it is easier to minimize the
augmented Lagrange function compared to solving the original problem
\eqref{eq:l1min} directly. In fact, the first step that minimizes $L_{\mu_k}
(\xx,\yy_k)$ assuming $\yy_k$ and $\mu_k$ are fixed can be solved element-wise
iteratively by a soft-thresholding algorithm \cite{WrightS2008,BeckA2009},
whose time complexity is bounded by $O(n^2)$ and can be easily parallelized.
Algorithm \ref{alg:alm} summarizes the ALM $\ell_1$-min algorithm \footnote{While we
present the ALM algorithm in the primal domain, there also
exists an implementation ALM in the dual domain.}.

@@

\section{Hardware Parallelism} \label{sec:parallelism} 
In this section we discuss the hardware architectures used in this paper. Aside
from the $\ell_1$-min algorithm, our choice of the architectures plays a
significant role in many practical decisions we make in Section
\ref{sec:alignment} and Section \ref{sec:recognition}.  For CPU implementation,
the system would make use of all the cores in as many CPUs are present on one
motherboard.  For GPU implementation, the system would make the best use of a
single many-core Nvidia GPU chip  on a PCI card (GPUs are commonly shipped with
a single GPU chip per PCI-express card).  Unless otherwise specified, all
implementations utilize single precision floating point datatypes.

\subsection{CPU Hardware Parallelism} 
The baseline architecture for our experiments is a Linux workstation with two
quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each processor
has its own memory interface, and is directly connected to half of the RAM
installed in the machine.  The amount of RAM installed exceeds the amount used
by the algorithms, and is not an important performance consideration.

Each core has a 32\,KB L1 data cache and a 256\,KB L2 cache. With Hyperthreading
enabled, each core can support two threads simultaneously.  For floating point
instructions, each core also has a vector unit capable of performing the same
arithmetic operation on four single precision floating point values
simultaneously.  \footnote{The core also has scalar floating point units, but
some compilers, notably gcc, generate code that uses the vector units
exclusively for floating point instructions.}

There are thus two important levels of parallelism that need to be exploited to
efficiently use a modern CPU: core-level parallelism and vector-level
parallelism.  For the implementations discussed here, we will primarily use the
OpenMP programming model to express core-level parallelism {\bf AY: What is
OpenMP? citation needed}.  Depending on the section of the code{\bf AY: What is
a section of the code?}, we will achieve vector-level parallelism implicitly
through either the auto-vectorization features of Intel's icc compiler, or
through the use of a highly optimized BLAS library (Intel's MKL). {\bf AY:
citations needed, what is icc, what is the relation between BLAS and MKL? Are
they equivalent? not clear in the context.}

Each processor further has 8MB of L3 cache that is shared by all four cores.
Overall, this means the algorithm has approximately 16MB of on-chip cache
available.  For the low resolution images used in sparse representation based
face recognition, this is enough cache to hold over 800 image-sized variables,
{\bf AY: how low resolution? which sparse representation problem? ``image-sized
variables'' means simply ``image vectors''?} meaning that for small data sets
it is possible to run completely using on-chip cache.

\subsection{GPU Hardware Parallelism}

The ALM algorithm requires many matrix-vector operations, this gives many-core
processors such as GPUs a computational advantage for large problem sizes.
GPUs are comprised of several streaming multiprocessors (SMP) (14 for our
NVIDIA Fermi) which each contains 32 floating-point units.  Each SMP contains
its own L1-cache and shared memory and all SMPs share a common L2-cache.  The
GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups
of 32 threads called \emph{warps}.  These warps are then grouped into higher
levels called \emph{thread blocks} and \emph{grids}.

Fermi's inverted memory hierarchy (aggregate l1 capacity is greater than l2 capacity) is designed to support massively parallel workloads in which there is more task-private data than shared data, providing an increase in performance and a computational advantage over CPUs.
For our system, each SMP has 64 kb of L1 cache with a total of 896 kb of L1 cache/shared memory and
768kb of L2 cache. However, this increase in performance does not come for
free. As the GPU has its own memory system, any data the GPU uses must be
transferred from RAM to the GPU over PCI-Express, which has limited bandwidth.
However, the cost of these transfers are amortized by doing large amounts of
computation at once, requiring code to be run completely on the GPU.  In 2011, NVIDIA is developing a new ARM-based CPU system, dubbed 
"Project Denver" to fix performance issues such as the limited bandwidth due to PCI-E \cite{Dally-Web}.

In summary, a programmable GPU can be analogous to a multi-core CPU processor,
except with more sophisticated vector processing hardware.  Each thread block
is a virtualized multiprocessor, which each thread as a virtualized scalar
processor.  All of these differences allow GPUs to perform intensive
calculations such as large scale matrix-vector operations at much faster rates than CPUs.

@@

The computation of the linearization is complicated by the need to avoid trivial 
solutions where the algorithm
progressively zooms in on a black pixel of the test image.
This is achieved by minimizing the $\ell_1$ error between the illumination model $Ax$ and the cropped,
resampled, and normalized test image $\tilde\bb$, rather than the un-normalized version as in 
conventional Lucas-Kanade iterative alignment.
This involves a bit of calculus:
\begin{equation}
\frac{\partial}{\partial \tau_p} \frac{\bb}{||\bb||} 
= \frac{1}{||\bb||} \frac{\partial \bb}{\partial \tau_p} - 
\frac{\bb}{||\bb||^3} \bb^T \frac{\partial \bb}{\partial \tau_p}
\end{equation}
If the warping $\tau$ is constrained to be a similarity transform,
it can be paramaterized in such a way that the computation of $\frac{\partial \bb}{\partial \tau_p}$
reduces to several linear vector operations involving pixel coordinates $(u,v)$, $\tau \in \Re^4$,
and the image gradients $\f_x, \f_y \in \Re^M$, the latter of which only has to be computed
once per test image. {bf DW: I could go into the rest of the math for this, but I'm worried about space}

The iterative alignment stage is summarized as Algorithm \ref{alg:iterative_alignment}.
\begin{algorithm}[thb]
\caption{\bf (Iterative Alignment Stage)} \label{alg:iterative_alignment}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} Aligned training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
$\f\in\Re^M$ 
\STATE $\tau^{(0)} \leftarrow \textrm{face detector($\f$)}$. 
\STATE $\f_x, \f_y \in \Re^M \leftarrow \frac{df}{dx}, \frac{df}{dy}$ 
\STATE {\bf for} each gallery subject $i$, 
\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
\STATE \hspace{6mm} $\tilde \bb(\tau) \leftarrow \frac{\bb \circ \tau}{\|\bb \circ \tau\|_2}$;
\STATE \hspace{6mm} $J \leftarrow  \frac{\partial}{\partial \tau} \tilde\bb(\tau)  \bigr|_{\tau^{(j)}} $;
\STATE \hspace{6mm} $\Delta \tau =  \arg\min \; \| \e \|_1  \; \subj \; \tilde \bb + J \Delta \tau = A_i \x + \e.$ 
\STATE \hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
\tau$; 
\end{small}
\end{algorithmic}
\end{algorithm}

The solution of the minimization problem in Algorithm
\ref{alg:iterative_alignment} Step 8 can be solved via a
variant of ALM, which is summarized as Algorithm \ref{alg:alignment_alm}.
$\gamma$ denotes the largest eigenvalue of the matrix $A^TA$. For the choice of
parameter $\mu$, we take the same strategy as in \cite{YangJ2009-pp} and set
$\mu_0 = 2m / \|\bb\|_1$. We set $\rho=1.5$.

\begin{algorithm}[h]
\caption{\bf (Augmented Lagrange Multiplier Method For Alignment)}
\begin{algorithmic}[1]
\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\blamda_1 = \mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\STATE $t_{1} \leftarrow 1, z_{1} \leftarrow \x_{k}, \u_{1} \leftarrow \x_{k}$;
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\e_{l+1} \leftarrow \textup{shrink}\left(\bb - A\x_l + \frac{\blamda_k}{\mu_k}, \frac{1}{\mu_k}\right)$;
\STATE $\x_{l+1} \leftarrow (A^\dagger)^T \left(\bb - \e_{l+1} + \frac{\blamda_k}{\mu_k} \right) $;
\ENDWHILE
\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
\STATE $\mu_{k+1} \leftarrow \rho\mu_k$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{algorithmic}
\label{alg:alignment_alm}
\end{algorithm}

\subsection{Computational Complexity} Any discussion of the speed of an
algorithm should discuss the comutational complexity of the algorithm.  For the
alignment stage ALM algorithm \ref{alg:alignment_alm}, this is rather simple.  The
Matrix-vector computions are $O(mn)$, and the vector-vector operations are
$O(m)$, in terms of both computation cost and memory-bandwidth.  Since A is
very tall, the computation of $A^\dagger$, the pseudoinverse of $A$, is best
computed via $A^\dagger = (A^TA)^{-1} A^T$.  The complexity of computing $G =
A^T A$ is $O(nnm)$.  $G$ is sufficiently small that it can be inverted via
Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation
of $A$ finishes with a matrix multiplication, with cost $O(n^2 m)$.
Fortunately, $A^\dagger$ only needs to be computed once per instance of the
$\ell_1$ minimization problem.

Therefore, for problems with large $n$, we expect the computation of the
pseudoinverse to dominate.  However, for the problem sizes needed in face
recognition problems, we find that the computation time is dominated by the
repeated computation of matrix-vector and matrix-matrix multiplications in the
inner loop.

%\STATE Compute residuals $r_i(\bb) = \| {\bb} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.

\subsection{CPU Implementation}

@@

%\subsection{Tuning the Soft-Thresholding Operator}
%...Make sure there are no branches
%...Make sure there are no function calls
%...Make sure the loop gets automatically vectorized
%...Block the computation into work for different cpu cores
\subsection{GPU Implementation}
{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}

@@

The first experiment is designed to compare the two implementations of PALM
that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \ref{alg:alm}) by
comparing the amount of time to solve a problem given a generic random matrix $A$ and a random sparse
vector $\xx_0$.  
The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
randomly generated from the standard normal and normalized with unit norm.  The ratio of $d:n$ varied from
$10:1$ to $10:5$ with the dimension of $d$ of matrix $A$ varying from 500 to 8000.
The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
Because the ground truth signal $\xx_0$ is known, the algorithm terminates
when $norm(\bx-\bx_0) < tolerance$.
The measurement vector was generated with the equation $\y_0 = A \x_0$.   
We also varied the different inner and outer tolerance levels from 1e-2 to 1e-5.
We averaged the the time taken for the ALM algorithm to solve 10 instances of a problem.

>>>>

\usepackage{iccv}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{amsfonts}
\usepackage{mydefs}
\usepackage{algorithm}
\usepackage{algorithmic}

@@

One of the drawbacks of most interior-point methods for $\ell_1$-min (including
BP, LASSO, and TNIPM) is that they require the solution sequence to follow an
interior path
%{\bf DW: Following the interior path is what made them faster than
%traditional linear programming techniques that follow the edge of the feasible set;
%does Homotopy actually go outside the feasible set?  Just want to check if this is
%an accurate representation...}
, which is often computationally expensive. To mitigate these
issues, an approach called \emph{Homotopy} has been studied to accelerate the
speed of $\ell_1$-min. Homotopy methods for recovering sparse signals were
first studied in the context of LASSO \cite{OsborneM2000}, which inspired a
solution to the \emph{forward stagewise linear regression} problem called LARS
\cite{EfronB2004}, and eventually led to the corresponding Homotopy algorithms
for BP in \cite{MalioutovD2005,DonohoD2006}.

@@

Note that in the iterative procedure outlined in \eqref{eq:alm}, the second
step only involves vector algebra and matrix-vector multiplication. Therefore,
the procedure is computationally efficient only if it is easier to minimize the
augmented Lagrange function compared to solving the original problem
\eqref{eq:l1min} directly. In fact, the first step that minimizes $L_{\mu_k}
(\xx,\yy_k)$ assuming $\yy_k$ and $\mu_k$ are fixed can be solved element-wise
iteratively by a soft-thresholding algorithm \cite{WrightS2008,BeckA2009},
whose time complexity is bounded by $O(n^2)$ and can be easily parallelized.
Algorithm \eqref{alg:alm} summarizes the ALM $\ell_1$-min algorithm \footnote{While we
present the ALM algorithm in the primal domain, there also
exists an implementation ALM in the dual domain.}.

@@

\section{Hardware Parallelism} \label{sec:parallelism} 
In this section we discuss the hardware architectures used in this paper. Aside
from the $\ell_1$-min algorithm, our choice of the architectures plays a
significant role in many practical decisions we make in Section
\ref{sec:alignment} and Section \ref{sec:recognition}.  For CPU implementation,
the system would make use of all the cores in as many CPUs are present on one
motherboard.  For GPU implementation, the system would make the best use of a
single many-core Nvidia GPU chip  on a PCI card (GPUs are commonly shipped with
a single GPU chip per PCI-express card).  Unless otherwise specified, all
implementations utilize single precision floating point datatypes.  Since on-chip
memory (cache) is often orders of magnitude faster than off-chip memory, the sizes
and speeds of on-chip memory (cache) have big performance implications.  We give
a brief overview of the caches that are available in our target architectures,
and defer discussion of their performance implications to sections
\ref{sec:alignment_implementation_cpu} and
\ref{sec:alignment_implementation_gpu}.

\subsection{CPU Hardware Parallelism} 
The baseline architecture for our experiments is a Linux workstation with two
quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each processor
has its own memory interface, and is directly interfaced to half of the RAM
installed in the machine.  The amount of RAM installed exceeds the amount used
by the algorithms, and is not an important performance consideration.

%2MB = 2**21 bytes
%64*64*4 = 2**14 bytes = 16384 bytes per image
%... so can fit 2**7  = 128 images.
Each core has a private 32\,KB L1 data cache and a private 256\,KB L2 cache.  
Between the 8 cores, this is a total of 2\,MB L2 cache.
For the low resolution images used in sparse representation based
face recognition (on the order of $64\times64$ pixel images), this corresponds to
128 images.

With Hyperthreading
enabled, each core can support two threads simultaneously.  For floating point
instructions, each core also has a vector unit capable of performing the same
arithmetic operation on four single precision floating point values
simultaneously.

%16MB = 2**24 bytes = 16777216 bytes
%64*64*4 = 2**14 bytes = 16384 bytes per image
%... so can fit 2**10  = 1024 images.
Each processor further has 8MB of L3 cache that is shared by all four cores.
Overall, this means the algorithm has approximately 16MB of on-chip cache
available.

it is possible
to simultaneously hold up to 1024 image vectors in on-chip cache at a time.

There are thus two important levels of parallelism that need to be exploited to
efficiently use a modern CPU: core-level parallelism and vector-level
parallelism. CPU tools for exploiting these two levels of parallelism are
relatively mature.  Heavily optimized Basic Linear Algebra Subprograms (BLAS) are
commercially available that make good use of both levels of CPU parallelism.
For operations that do not fit well with the BLAS API, we 
achieve both vector-level and core-level parallelism through a combination of
the OpenMP API \cite{dagum2002openmp} and the automatic vectorization 
facilities of Intel's C++ compiler (ICC)\cite{dulong1999overview}.
% DW: I'm really not sure these citations are even really relevant;
% our target audience will just google icc openmp and get more up-to-date
% info anyway.
\subsection{GPU Hardware Parallelism}
Compared to CPU architectures, GPU architectures dedicate less of their transistors
to cache, and more of them for arithmetic logic. Our target GPU architecture
is the Nvidia Fermi architecture.  CUDA based GPUs are comprised of several 
streaming multiprocessors (SMP) (currently 14 to 16 for a high end GPU).

The GPU executes programs as SIMT (Single Instruction Multiple Threads) in
groups of 32 threads called \emph{warps}.  These warps are then grouped into
higher levels called \emph{thread blocks} and \emph{grids}.  It is impossible
to communicate between SMPs during gpu program execution, computational tasks
are typically broken up into multiple gpu programs (kernels) that are executed
serially.

Each SMP contains its own L1-cache, which is divided between hardware managed
and software managed ("shared") memory.  Additionally, all SMPs share a common
L2-cache.  For our system, each SMP has 64 kb of L1 cache with a total of 896
kb of L1 cache/shared memory and 768kb of L2 cache. The relatively small
amount of cache on the GPU is ballanced by a significantly higher bandwidth between the
processor chip and off-chip memory (DRAM) compared to the CPU.

The GPU has its own memory system, any data the GPU uses must first be
transferred from CPU DRAM to GPU DRAM over PCI-Express.  For our application,
this transfer overhead is amortized over a large amount of computation and is
not a major concern.

Compared to a typical CPU, a GPU has much higher peak computational bandwidth
as well as data bandwith to DRAM.  However, since our application is primarily
of matrix-vector operations and image resampling operations, it has a very low
arithmetic intensity (i.e. we perform a very small number of arithmetic
operations for every number that has to be loaded), and thus performance is
typically constrained by memory (and cache) bandwidth.  At large problem sizes
the CPU is limited by its relatively low bandwidth to DRAM compared to the gpu,
while at small problem sizes the GPU is limited by the relatively high latency
of its access to DRAM.

@@

The computation of the linearization is complicated by the need to avoid trivial 
solutions where the algorithm
progressively zooms in on a black pixel of the test image.
This is achieved by minimizing the $\ell_1$ error between the illumination model $Ax$ and the cropped,
resampled, and normalized test image $\tilde\bb$, rather than the un-normalized version as in 
conventional Lucas-Kanade iterative alignment.
This involves a bit of calculus:
\begin{equation}
\frac{\partial}{\partial \tau_p} \frac{\bb}{||\bb||} 
= \frac{1}{||\bb||} \frac{\partial \bb}{\partial \tau_p} - 
\frac{\bb}{||\bb||^3} \bb^T \frac{\partial \bb}{\partial \tau_p}
\end{equation}
If the warping $\tau$ is constrained to be a similarity transform,
it can be paramaterized in such a way that the computation of $\frac{\partial \bb}{\partial \tau_p}$
reduces to several linear vector operations involving pixel coordinates $(u,v)$, $\tau \in \Re^4$,
and the image gradients $\f_x, \f_y \in \Re^M$, the latter of which only has to be computed
once per test image.  The iterative alignment stage is summarized as Algorithm \eqref{alg:iterative_alignment}.
\begin{algorithm}[thb]
\caption{\bf (Iterative Alignment Stage)} \label{alg:iterative_alignment}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} Aligned training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
$\f\in\Re^M$ 
%\STATE $\tau^{(0)} \leftarrow \textrm{face detector($\f$)}$. 
\STATE $\f_x, \f_y \in \Re^M \leftarrow \frac{df}{dx}, \frac{df}{dy}$ 
\STATE {\bf for} each gallery subject $i$, 
\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do} 
\STATE \hspace{6mm} $\tilde \bb(\tau) \leftarrow \frac{\bb \circ \tau}{\|\bb \circ \tau\|_2}$;
\STATE \hspace{6mm} $J \leftarrow  \frac{\partial}{\partial \tau} \tilde\bb(\tau)  \bigr|_{\tau^{(j)}} $;
\STATE \hspace{6mm} $\Delta \tau =  \arg\min \; \| \e \|_1  \; \subj \; \tilde \bb + J \Delta \tau = A_i \x + \e.$ 
\STATE \hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
\tau$; 
\end{small}
\end{algorithmic}
\end{algorithm}

The solution of the minimization problem in Algorithm
\eqref{alg:iterative_alignment} Step 7 can be solved via a
variant of ALM, which is summarized as Algorithm \eqref{alg:alignment_alm}.
$\gamma$ denotes the largest eigenvalue of the matrix $A^TA$. For the choice of
parameter $\mu$, we take the same strategy as in \cite{YangJ2009-pp} and set
$\mu_0 = 2m / \|\bb\|_1$. We set $\rho=1.5$.

\begin{algorithm}[h]
\caption{\bf (Augmented Lagrange Multiplier Method For Alignment)}
\begin{algorithmic}[1]
\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\blamda_1 = \mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\STATE $t_{1} \leftarrow 1, z_{1} \leftarrow \x_{k}, \u_{1} \leftarrow \x_{k}$;
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\e_{l+1} \leftarrow \textup{shrink}\left(\bb - A\x_l + \frac{\blamda_k}{\mu_k}, \frac{1}{\mu_k}\right)$;
\STATE $\x_{l+1} \leftarrow (A^\dagger)^T \left(\bb - \e_{l+1} + \frac{\blamda_k}{\mu_k} \right) $;
\ENDWHILE
\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
\STATE $\mu_{k+1} \leftarrow \rho\mu_k$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{algorithmic}
\label{alg:alignment_alm}
\end{algorithm}
%\STATE Compute residuals $r_i(\bb) = \| {\bb} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.

\subsection{Computational Complexity} Any discussion of the speed of an
algorithm should discuss the comutational complexity of the algorithm.  For the
alignment stage ALM algorithm \eqref{alg:alignment_alm}, this is rather simple.
The Matrix-vector computions are $O(mn)$, and the vector-vector operations are
$O(m)$, in terms of both computation cost and memory-bandwidth.  Since A is
very tall, the computation of $A^\dagger$, the pseudoinverse of $A$, is best
computed via $A^\dagger = (A^TA)^{-1} A^T$.  The complexity of computing $G =
A^T A$ is $O(nnm)$.  $G$ is sufficiently small that it can be inverted via
Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation
of $A$ finishes with a matrix multiplication, with cost $O(n^2 m)$.
Fortunately, $A^\dagger$ only needs to be computed once per instance of the
$\ell_1$ minimization problem.

Therefore, for problems with large $n$, we expect the computation of the
pseudoinverse to dominate.  However, for the problem sizes needed in face
recognition problems, we find that the computation time is dominated by the
bandwidth required by repeated computation of matrix-vector and matrix-matrix
multiplications in the inner loop.

I summary, the alignment stage essentially two levels of available parallelism:
at the higher level there are multiple alignent problems that are solved
independently.  At a lower level, the first-order linear algebraic operations
exhibit parallelism within image operations, i.e. at a pixel level.

\subsection{CPU Implementation} 
\label{sec:alignment_implementation_cpu}

This section discusses how the iterative
alignment problem can be implemented on CPU hardware.  The implementations
presented in \cite{WagnerA2009-CVPR} and \cite{WagnerA2011-PAMI} were written
using the OpenCV library.

OpenCV wraps either the operating system's built-in BLAS library, or Math
Kernel Library (MKL), Intel's BLAS implementation, and thus achieves vector
(i.e. pixel)-level parallelism implicitly.  The authors of
\cite{WagnerA2011-PAMI} used OpenMP to schedule instances of the alignment
problem to the CPU cores in a 1:1 fashion.  With eight alignment problems for
ten training users being solved concurrently, and with 20-40 images per
training user, all of the required data fits into L3 cache.  This provided a
significant, though sub-linear improvement over a single-core implementation.

In the proposed CPU implementation, we instead solve a single alignment
instance at a time, using a multi-core BLAS library.  Recall that the L2 cache
can old up to 128 image vectors' worth of data.  Since we're only aligning data
for 1 gallery user at a time the working set of the problem fits into L2 cache,
resulting in significantly higher image alignment speed than previous cpu
implementations.

@@

%\subsection{Tuning the Soft-Thresholding Operator}
%...Make sure there are no branches
%...Make sure there are no function calls
%...Make sure the loop gets automatically vectorized
%...Block the computation into work for different cpu cores
\subsection{GPU Implementation}
\label{sec:alignment_implementation_gpu}
{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}

@@

The first experiment is designed to compare the two implementations of PALM
that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \eqref{alg:alm}) by
comparing the amount of time to solve a problem given a generic random matrix $A$ and a random sparse
vector $\xx_0$.  
The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
randomly generated from the standard normal and normalized with unit norm.  The ratio of $d:n$ varied from
$10:1$ to $10:5$ with the dimension of $d$ of matrix $A$ varying from 500 to 8000.
The ground truth signals, $\xx_0$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
Because the ground truth signal $\xx_0$ is known, the algorithm terminates
when $norm(\bx-\bx_0) < tolerance$.
The measurement vector was generated with the equation $\y_0 = A \x_0$.   
We also varied the different inner and outer tolerance levels from 1e-2 to 1e-5.
We averaged the the time taken for the ALM algorithm to solve 10 instances of a problem.
